I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-12-07 paper_io_sim.tex
[WIP] intro fixes

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{ulem}
\usepackage{tabularx}

% Fix link colors
\hypersetup{
    colorlinks = true,
    linkcolor=red,
    citecolor=red,
    urlcolor=blue,
    linktocpage % so that page numbers are clickable in toc
}

@@

        The rapid emergence of Big Data seen in recent years has resulted in a 
        growing needs for efficient data storing and optimized data analysis. 
        Despite relying on large computing infrastructures with powerful computing capability 
        to process these large datasets, an I/O bottleneck remains. 
        One widely-used approach to minimize the costs of data transfers during
        regular use is through the employment of the page cache.
        While many simulation tools have been developed to 
        support studies in computing systems optimization, page cache simulation 
        has not been fully captured or remains absent from these simulators. 
        The lack of ability to simulate I/O with page cache leads to inaccurate 
        simulation results in many situations, especially when dealing with data intensive applications.

        In this paper, we propose a I/O simulation model that characterizes the key features of the page cache.
        We have implemented the model with SimGrid, a simulation framework for 
        distributed systems, using WRENCH, a SimGrid-based workflow simulator. 
        Our simulator is compatible with both single-threaded and multithreaded applications,
        and can simulate both writeback and writethrough cache that can factor into local
        and network-based filesystems.
        The results show that our simulator not only achieves higher accuracy than 
        original WRENCH, but also effectively simulates memory contents and 
        page cache during I/O operations. 
        Through this study, our effort is to contribute an analytical model 
        of the page cache, which can also be integrated to other simulators 
        to improve I/O simulation accuracy, with its practical results.

@@

        In recent years, as a result of the growing demand for Big Data processing 
        on large computing infrastructures, there have been the research and development 
        of simulation applications and frameworks for distributed systems. 
        These simulation tools have shown their advantages in evaluation 
        of algorithms, applications and solutions. 
        Experimenting on real platforms exposes shortcomings due to limited and 
        shared resources, operational costs, unstable environments and the result 
        reproducibility difficulties.
        On the contrary, simulators provide to researchers the freedom of choosing 
        simulated infrastructures, reproducible results and low costs.
        These simulators are developed and integrated with simulation models 
        of computing resources including CPU, network and storage.
        The ultimate goal that all simulators try to achieve is to accurately simulate 
        applications. 
        However, there is a trade-off between accuracy and scalability, which has been 
        indicated in some reviews \cite{casanova2014simgrid} \cite{byrne2017review}. 
        This means that in most ciscumstances, users have to, more or less, sacrifice 
        accuracy for the scalability and performance when their platforms grow to 
        hundreds or thousands nodes.

        Beside computing resources such as CPU, network, storage, another 
        important factor that has significant impacts on I/O performance is 
        page cache. 
        Page cache is an architectural solution employed in systems to tackle 
        the problem of disk I/O bottleneck in data processing. 
        Although this is not a radical solution for disk performance, it contributes 
        to minimizing disk accesses and maximizing I/O performance. 
        Surprisingly, page cache has not been fully modeled, if not completely absent, 
        in simulation tools. 
        Some simulators try to simulate I/O with page cache, but fail to capture the 
        key features such as dirty data and cache eviction mechanism 
        \cite{nunez2012simcan} \cite{nunez2012icancloud}.
        Again, these simulators also have a problem in scalability with 
        long simulation time regarding to their very detailed level of simulation. 
        Other works are able to capture important features of page cache, 
        but they focus on very specific domains, such as energy consumption 
        in \cite{xu2018saving}.

        In this work, we propose a simulation model of I/O with page cache, 
        implement it with SimGrid, a simulation framework for distributed computing, 
        and WRENCH, an open-source library for developing workflow simulator 
        built on top of SimGrid. 
        One of the reasons for us to choose SimGrid is that it has a balance 
        between accuracy and scalability. Moreover, the versatility of SimGrid 
        enables it to be used in cross-domain \cite{casanova2014simgrid}. 
        In addition, SimGrid has been actively maintained with updated documents. 
        Then we evaluate our model with four simulators in four experiment scenarios. 
        This paper is organized as follows. 
        In Section \ref{relatedwork} we provide details of page cache 
        existing simulation tools.
        In Section \ref{relatedwork} we propose our I/O with page 
        cache simulation model and how we implement it and our experiments 
        to evaluate the model.  
        Section \ref{results} compares the results of our model with the results 
        of original WRENCH and real execution results. 
        Finally, Section \ref{discussion} concludes and presents future 
        work.

@@

    \section{Method}
    \label{method}
    In this section, we present our page cache simulation model for file I/O,
    flusing, and eviction. We detail the design and implementation of two simulators:
    (1) a standalone Python prototype to evaluate the simulation model independently of any other 
    tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
    We also describe experimental scenarios to evaluate our simulators in various configurations.

    We separate our simulation model in two components, the IO
    Controller and the Memory Manager, which together simulate 
    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
    To read or write a file chunk, a simulated application sends a
    request to the IOController. The IOController interacts as needed with
    the Memory Manager to free memory through flushing or eviction,
    and to read or write to cache. The Memory Manager
    implements these operations, simulates periodical flushing
    and eviction from cache, and reads or writes to disk when necessary.
    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the simulated writeback page cache.
           Applications send file read or write requests to the
           IOController that orchestrates flushing, eviction, cache
           and disk accesses with the MemoryManager. Storage
           devices (memory and disk) simulate file transfers with
           bandwidth sharing.}
           \label{fig:interaction}
    \end{figure}

    \subsection{Memory Manager}

    The Memory Manager simulates two parallel threads: the main one
    implements flushing, eviction, and cache I/Os synchronously, whereas
    the second one, operating in the background, periodically searches for
    expired dirty data in page cache LRU lists and flushes them to disk. We
    use existing storage simulation models~\cite{lebre2015} to simulate
    memory, characterized by its storage capacity, read and write
    bandwidths, and latency. Reusing existing storage models allows us to
    simulate bandwidth sharing between concurrent memory accesses.

    \subsubsection{Page cache LRU lists}

    In the Linux kernel, page cache LRU lists contain file pages. However, 
    due to the large number of file pages, managing lists of pages 
    induces substantial overheads.
    Therefore, we introduce the concept of a data block as a unit to represent data 
    cached in memory. A data block is a subset of file pages stored in
    page cache that were accessed in the same I/O operation. 
    A data block has information about file name, block size, last access 
    time, a dirty flag that represents whether the data is clean (0) 
    or dirty (1), and an entry (creation) time.
    Blocks can have different sizes and a given file can have multiple 
    data blocks in page cache. In addition, a data block can be split into an 
    arbitrary number of smaller blocks.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
           \caption{Model of page cache LRU lists with data blocks.}    \label{fig:lrulist}
    \end{figure}

    We model page cache LRU lists as 
    two lists of data blocks, an active list and an inactive list, both ordered by 
    last access time (earliest first, Figure~\ref{fig:lrulist}).
    As in the kernel, our simulator limits the size of the active list to
    twice the size of the inactive list, by moving least recently 
    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
    % Modeling page cache LRU lists as lists of data blocks reduces the
    % overhead of the simulator, as data blocks are obviously less
    % numerous than file pages. At the same time, it achieves the accuracy in 
    % I/O time simulation that can be seen in the later section.
    % From Tristan: I don't think the sentences above are necessary,
    % they are quite repetitive.

    At any given time, a file can be partially cached, completely cached,
    or uncached at all. Cached data blocks may reside in one or both of the
    page cache LRU lists. The first time they are accessed, blocks are
    added to the inactive list. On subsequent accesses, blocks of the
    inactive list are moved to \tristan{if they are moved then they can't
    be in both lists} the top of the active list. Cached blocks written to
    cache are marked as dirty until they are flushed to disk.

    \subsubsection{Reads and Writes}

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
    accessed in a round-robin fashion rather than fully randomly. 
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created 
    and put at the top of the inactive list. 
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
    and put it in the active list. 
    If the blocks are dirty, we move them to the active list to preserve their entry time. 
    Because the chunk size and block size can be different, there are situations
    where a block is not entirely read. 
    In this case, the block is split in two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data 
           is read first, followed by data from the inactive list, and finally data from the active list.}
           \label{fig:read_order}
    \end{figure}

    For file writes, we assume that all data to be written is 
    uncached. Thus, each time a chunk is written, we create a block of dirty data 
    and add it to the inactive list.

    \subsubsection{Flushing and eviction}

    The main simulated thread in the Memory Manager can flush the
    cache to write dirty data to disk. The data flushing simulation
    function takes the amount of data to flush as parameter. While
    the amount of data to flush is not reached and there are dirty
    blocks remaining in cache, this function traverses the sorted
    inactive list, then the sorted active list, and writes the
    least recently used dirty block to disk having set its dirty
    flag to 0. In case the amount of data to flush requires that a
    block be partially flushed, the block is split in two blocks,
    one that is flushed and one that remains dirty. Flushing time
    to disk is simulated with the storage model.

    Similarly to data flushing, our cache eviction simulation runs in
    the main thread. It frees up the page cache by traversing and deleting 
    least recently used clean data blocks in the inactive list.
    The amount of data to evict is passed as a parameter and data blocks are deleted 
    from the inactive list until total evicted data reaches the required amount,
    or until there is no clean block left in the list.
    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
    and only one of them is evicted.
    The cache eviction simulation does not contribute to the simulated time 
    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.

    \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{in}{page cache inactive list}
                \Desc{ac}{page cache active list}
                \Desc{t}{predefined flushing time interval}
                \Desc{exp}{predefined expiration time}
                \Desc{sm}{storage simulation model}
               \EndInput
               \While{host is on}
                \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                \State flushing\_time = sm.write(blocks)
                \State flag blocks as clean data \tristan{a bit informal}
                \If{flushing\_time $<$ t}
                    \State sleep(t - flushing\_time)
                \EndIf \tristan{else?}
            \EndWhile
        \end{algorithmic}
    \end{algorithm}

    Periodical flushing is simulated in the Memory Manager
    background thread. As in the Linux kernel, a dirty block
    in our model is considered expired if (1) it is dirty, and (2)
    the duration since its entry time is longer than a
    predefined expiration time. 
    Periodical flushing is simulated as an infinite loop in which 
    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}). 
    % From Tristan: the algorithm is quite straightforward, I don't think this is 
    % necessary
    % In each repetition, Memory Manager finds expired dirty blocks in two 
    % page cache LRU lists (line 8), simulates writes of data of these blocks 
    % to disk (line 9), and mark them as clean (line 10).
    % If the flushing time does not exceed our time interval, the thread is put 
    % to sleep for the remaining time (lines 11-13). 
    % Then based on the state of the host at current simulated time, 
    % the algorithm continues or finishes the loop.
    Because periodical flushing is simulated as a background thread, it can happen concurrently
    with disk I/O initiated by the main thread. This parallelism is handled by the 
    storage model and reflected in simulated I/O time.

    \subsection{I/O Controller}

    \begin{algorithm}\caption{File chunk read simulation of IOController}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn))
               \State cache\_read = cs - disk\_read
            \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
            \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
            \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
                \State sm.read(disk\_read)  
                \State mm.add\_to\_cache(disk\_read, fn)     
            \EndIf
            \If {cache\_read $>$ 0} \Comment{Read cached}
                \State mm.cache\_read(cache\_read)  
            \EndIf
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IOController which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 7-8). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache \tristan{Create a variable
    "required\_mem" in the algorithm to clarify}.
    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). When called with negative arguments, functions
    flush and evict simply return and do not do anything. Then, if there is
    uncached data, a disk read is simulated with disk model (line 12)
    \tristan{this is not consistent with the new Figure 1 where disk reads
    are done by the mm}, and this data of the file is added to cache (line
    13).
    Finally, if cached data needs to be read, the Memory Manager is called
    to simulate a cache read  
    and update the corresponding data blocks accordingly (line 16).

    \begin{algorithm}\caption{File chunk write simulation of IOController}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write(fn, mem\_amt) 
            \EndIf
            \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)  
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IOController is not supposed to interact directly with sm}
            \EndIf

        \end{algorithmic}
    \end{algorithm}
    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
    the IOController. 
    % Data is written to cache until the amount of dirty data in the cache 
    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
    Our algorithm initially checks the  amount of data that 
    can be written as dirty data given the dirty ratio (line 6).
    If this amount is greater than 0, the Memory Manager is requested to evict 
    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained somewhere}.
    After eviction, the amount of data that can be written to 
    page cache is calculated (line 9), and a cache write is simulated (line 10).
    To write the remaining data, we flush and 
    evict data as possible so that this data can be 
    written to cache (line 13-14). 
    The Memory Manager does not guarantee that the amount of flushed 
    and evicted data is enough to write the remaining data. 
    In the case it is less than the required amount, the remaining data 
    being written to cache can be flushed and evicted right away. 
    Thus, the amount of remaining data written and kept in cache 
    after the write is limited by the amount free memory, and this amount 
    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
    Finally, the disk write is simulated with the remaining data amount (line 16), 
    and the chunk write algorithm finishes \tristan{see comment in algorithm}.

@@

            Since bandwidth sharing was not simulated, this prototype does not support
            concurrency: it is limited to single-threaded applications running on systems
            with a single-core CPU. We used this prototype for a first validation of our simulation
            model against a real sequential application running on a real system.
            The source code this prototype is available at
            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}. We used Python 3.7.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{ulem}

@@

		The rapid emergence of Big Data seen in recent years has resulted in a 
		growing need for efficient data storing and optimized data analysis techniques. 
		Despite relying on large computing infrastructures with powerful computing capability 
		to process these large datasets, an I/O bottleneck remains. 
        One widely-used approach to minimize the costs of data transfers during
        regular use is through the employment of the page cache.
		While many simulation tools have been developed to 
		support studies in computing systems optimization, page cache simulation 
		has not been fully captured or remains absent from these simulators. 
		The inability to simulate I/O with page cache leads to inaccurate 
		simulations in many situations, especially when dealing with data intensive applications.

		In this paper, we propose a I/O simulation model that characterizes the key features of the page cache.
		We have implemented the model with SimGrid, a simulation framework for 
		distributed systems, using WRENCH, a SimGrid-based workflow simulator. 
        Our simulator is compatible with both single-threaded and multithreaded applications,
        and can simulate both writeback and writethrough cache that can factor into local
        and network-based filesystems.
		The results show that our simulator not only achieves higher accuracy than 
		original WRENCH, but also effectively simulates memory contents and 
		page cache during I/O operations. 
		Through this study, our effort is to contribute an analytical model 
		of the page cache, which can also be integrated to other simulators 
		to improve I/O simulation accuracy.

@@

        The Linux Page Cache, introduced in kernel version 2.4, plays an important role in
        the reduction of filesystem data transfer costs~\cite{pagecache}. With the page cache, previously
        read data can be reread directly from memory and written data can be written to
        memory before being asynchronously flushed to disk, resulting in improved I/O performance
        on slower storage devices. The performance benefits that can be obtained from using the page
        cache are determined by many factors including the total amount of memory,
        the amount of data being written (i.e. dirty data), and the percentage of available memory available for
        written data. All these factors are important when determining the impact of data transfers on
        application performance, particularly in data intensive applications.

        Characterizing the performance of applications directly on computing infrastructure
        has various shortcomings, which include, limited and shared resources, operational
        costs, unstable environments, and the difficulties in producing reproducible results.
        Simulations address these concerns by providing abstractions to the performance of computer
        hardware, such as CPU, network and storage. As a result, simulations provide and
        cost-effective, fast, easy and reproducible  way evaluate application performance
        on different infrastructures without having to execute the code.

        Page cache simulation is crucial when simulating data intensive applications,
        which would rely on cache hits and writing to memory for additional performance.
        While existing simulators can adequately characterize many different computing environments,
        they lack much of the ability to adequately describe the page cache. Some
        simulators try to simulator I/O with page cache, but fail to capture the key features such
        as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
        Other simulators, such as~\cite{xu2018}, capture the important feature of the
        page cache, but are domain-specific.

		%However, there is a trade-off between accuracy and scalability, which has been 
		%indicated in some reviews \cite{casanova2014simgrid} \cite{byrne2017review}. 
		%This means that in most cicumstances, users have to, more or less, sacrifice 
		%accuracy for the scalability and performance when their platforms grow to 
		%hundreds or thousands nodes.

        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
        cache incorporated into WRENCH, a workflow management system simulator based of
        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
        with an accurate representation of a workflow execution, when file system page
        cache is used. WRENCH-Cache is able to simulate both writeback and writethrough
        cache which would normally be available to local and NFS-based filesystems.

@@

	\section{Method}
	\label{method}

		In this section, we present our approach to model file read and write with 
		page cache, cache eviction and data flushing mechanisms implemented 
		in the Linux kernel. We also detail the design and implementation of two simulators:
		(1) a standalone Python prototype used to evaluate the model independently of any other 
		library, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
		Finally, we describe experimental scenarios to evaluate our simulations in various configurations.

		\subsection{Principle of the simulator}

			We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read and write mechanisms in Linux (Fig.~\ref{fig:interaction}). 
			To read or write a file chunk, a simulated application sends a request 
			to the IOController. The IOController determines the amount of memory required by the operation,
			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
			and finally reads or writes on disk if necessary.
			The Memory Manager simulates flushing, eviction, and I/O to cache.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
				   \caption{Overview of the simulated writeback page cache.
				   Applications send file read or write requests to the
				   IOController that orchestrates flushing, eviction, cache
				   and disk accesses with the MemoryManager. Storage
				   devices (memory and disk) simulate file transfers with
				   bandwidth sharing. \tristan{Could you annotate arrows to
				   make them more explicit?}.}
				   \label{fig:interaction}
			\end{figure}

			\subsubsection{Memory simulation model}

			We leverage existing storage simulation models~\cite{lebre2015} to simulate the memory,
			characterized by its storage capacity, read and write bandwidths,
			and device latency.
			Leveraging existing storage models allows us 
			to take advantage of bandwidth sharing between concurrent I/Os to 
			simulate parallel memory accesses.

			\paragraph*{Memory Manager}
			The Memory Manager simulates  two parallel threads: the first and main one simulates 
			flushing, eviction, and cache I/Os synchronously; the second one is a background thread that  
			periodically searches for expired dirty data in page cache LRU lists 
			and flushes them to disk.

			\paragraph*{Page cache LRU lists}
			In the Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages 
			induces substantial overheads.
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
   				\caption{Model of page cache LRU lists with data blocks.}	\label{fig:lrulist}
			\end{figure}

			We model page cache LRU lists as 
			two lists of data blocks, an active list and an inactive list, both ordered by 
			last access time (earliest first, Fig.~\ref{fig:lrulist}).
			As in the kernel, our simulator limits the size of the active list to
			twice the size of the inactive list, by moving least recently 
            used data blocks from the active list to the inactive list.
			Modeling page cache LRU lists as lists of data blocks reduces the
			overhead of the simulator, as data blocks are obviously less
			numerous than file pages, and at the same time achieves the accuracy in 
			I/O time simulation that can be seen in the later section.

			At a specific time, an arbitrary file can be partially cached, completely cached, 
			or uncached.
			Cached data is composed of data blocks existing in one or both
            of the page cache LRU lists. 
            Initially, blocks accessed for the first time are added to the inactive list in 
            page cache. 
            If the blocks residing the inactive list are subsequently accessed, 
            they will be moved to the top of the active list
            Cached blocks written to cache are marked as dirty blocks until 
            they are flushed to disk.

			\paragraph*{File read and write}
            Our model supports chunk-by-chunk file read and write simulation
            with a user-defined chunk size, in which chunks are read/written
            until file is entirely read/written.
            The simulated LRU lists are used as follows. 
            When a file is read, its data can be partially cached, completely cached, 
            or uncached. 
            In the scope of this study, our model support file accesses in chunk 
            but not file random access.
            Thus, we simplify the data read order, in which uncached data is read 
            from disk first, and cached data is only read after all uncached data is read 
            (left to right in Fig.~\ref{fig:read_order}).
            When a chunk of uncached data is read, a new clean block is created 
            and put at the top of the inactive list. 
            For a chunk of cached data, one or more existing data blocks in two 
            page cache LRU lists are accessed again, and moved to the active list. 
			If the blocks are clean, we only need to keep the information about the 
			amount of data as they are available for eviction. 
			Thus, these re-accessed clean blocks can be merge into a bigger block 
			to reduce the number of block objects and put it in the active list. 
            If the blocks are dirty, since the blocks need to keep the information 
            of entry time for periodical flushing, they are only moved to the active list. 
			Because the chunk size and block size can be different, there are circumstances 
			in which a block is not entirely read. 
			In this case, the block is split into two smaller blocks, one is re-accessed 
			and the other remains the same.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
   				\caption{File data read order from left to right, in which uncached data 
   				is read first, then cached data from inactive list and active list is read later.}	
   				\label{fig:read_order}
			\end{figure}

			For writes, we assume that all data to be written is 
			uncached and it is written to cache as dirty data. 
			Thus, each time a chunk is written, we create a block of dirty data 
			and put it in the inactive list.

            As is done in the kernel, we ensure the size of the active list does not 
            exceed twice the size of the inactive list 
			\cite{gorman2004understanding, linuxdev3rd2010}. 
			This is done by moving least recently used blocks in the active list 
			to the inactive list when the active list grows. If only a part of a data 
			block needs to be moved, the block is split.

			\paragraph*{Flushing}
			Next, we simulate data flushing, which runs in the main simulated 
			thread and writes back dirty data to disk. 
			The data flushing simulation function traverses the sorted 
			inactive list first, and then the active list, looks for least recently used 
			dirty blocks, writes them to disk through the storage device model, 
			and sets dirty flags of these blocks to 0 until the total flushed amount 
			reaches the amount to be flushed or there is no dirty data left in cache. 
			Similar to file read and write, the size of flushed data and block size 
			can be different, there are cases that the last block is not entirely flushed.  
			In that case, the last block is split into two blocks, one that is flushed 
			and one that remains dirty.
			The flushing time is simulated with storage model.

			\paragraph*{Cache eviction}	
			Similarly to data flushing, our cache eviction simulation runs on 
			the main thread, frees up the page cache by traversing and deleting 
			least recently used clean data blocks in the inactive list.
			The amount of evicted data is passed in and data blocks are deleted 
			from the inactive list until total evicted data reaches the amount 
			passed in, or there is no clean block left in the list.
			If the last evicted block is not entirely evicted, the block is split, 
			and only one block is deleted.
			The cache eviction simulation does not add up to the simulated time 
			since cache eviction time is negligible in real systems.

			\begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{in}{page cache inactive list}
        				\Desc{ac}{page cache active list}
						\Desc{t}{predefined flushing time interval}
						\Desc{exp}{predefined expiration time}						
						\Desc{sm}{storage simulation model}
   					\EndInput
   					\While{host is on}
						\State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
						\State flushing\_time = sm.write(blocks)
						\State flag blocks as clean data
						\If{flushing\_time $<$ t}
							\State sleep(t - flushing\_time)
						\EndIf
					\EndWhile
				\end{algorithmic}
			\end{algorithm}

			\paragraph*{Periodical flushing}	
			The periodical flushing simulation is run on a simulated background thread 
			to flush expired dirty data blocks from cache to disk.
			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
			is considered expired if: 
			\begin{enumerate}
				\item It is dirty and 
				\item The time from its entry time to current time is longer than the 
				predefined expiration time. 
			\end{enumerate}
			The simulation algorithm of periodical flushing is an infinite loop in which 
			Memory Manager searches and flushes these expired dirty blocks to disk, 
			as described in Algorithm~\ref{alg:pdflush}. 
			The loop ends when the host is off (line 7). 
			In each repetition, Memory Manager finds expired dirty blocks in two 
			page cache LRU lists (line 8), simulates writes of data of these blocks 
			to disk (line 9), and mark them as clean (line 10).
			If the flushing time does not exceed our time interval, the thread is put 
			to sleep for the remaining time (lines 11-13). 
			Then based on the state of the host at current simulated time, 
			the algorithm continues or finishes the loop.
			Because periodical flushing is simulated as a background thread, it can be 
			concurrent with disk I/O on main thread. This parallelism is handled by the 
			storage model and reflected in simulated I/O time.

			\subsubsection{File I/O simulation model}

			As mentioned, in our model, file reads and writes can be completed by 
			reading or writing chunk-by-chunk until entire file is read or written.

			\begin{algorithm}\caption{File chunk read simulation of IOController}
			\label{alg:read}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{cs}{chunk size}
        				\Desc{fn}{file name}
        				\Desc{fs}{file size (assume to fit in memory)}
						\Desc{mm}{MemoryManager object}
						\Desc{sm}{storage simulation model}
   					\EndInput
   					\State disk\_read = min(cs, fs - mm.cached(fn))
   					\State cache\_read = cs - disk\_read
					\State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
					\State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
					\If {disk\_read $>$ 0}  \Comment{Read uncached data}	
						\State sm.read(disk\_read)  
						\State mm.add\_to\_cache(disk\_read, fn) 	
					\EndIf
					\If {cache\_read $>$ 0} \Comment{Read cached}
						\State mm.cache\_read(cache\_read)  
					\EndIf
				\end{algorithmic}
			\end{algorithm}

			When a simulated application reads a file chunk, it sends a chunk read request 
			to IOController. 			
			The algorithm to simulate a chunk read implemented in IOController 
			is described in Algorithm~\ref{alg:read}.   
			First, we calculate the amount of uncached data that needs to be read 
			from disk, and the remaining amount is read from cache (line 7-8).
			We assume that after a file is read, there are two copies of the file in memory 
			(one in cache, one in anonymous memory). 
			Unfortunately, SimGrid lacks of application memory management and WRENCH 
			only use the memory capacity to check if the applications can be run. 
			Thus, in our model, when a chunk is read, one copy of the chunk 
			occupies anonymous memory, and the uncached part of that chunk is 
			read into cache.	
			The amount of required memory to read the chunk is calculated and 
			if there is not enough memory available, Memory Manager is called to 
			flush dirty data to disk and simulate flushing time with memory model (line 9). 
			This flushing is complemented by eviction in the Memory Manager (line 10). 
			Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
			and this data of the file is added to cache (line 13).
			Finally, if cached data is read, Memory Manager is called to simulate a cache read  
			and update the corresponding data blocks in cache (line 16).

			\begin{algorithm}\caption{File chunk write simulation of IOController}
			\label{alg:write}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{cs}{chunk size}
        				\Desc{fn}{file name}
						\Desc{mm}{MemoryManager object}
						\Desc{sm}{storage simulation model}
   					\EndInput
					\State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
					\If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
    					\State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
    					\State mem\_amt = min(cs, mm.free\_mem)
    					\State mm.write(fn, mem\_amt) 
    				\EndIf
					\If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
						\State mm.flush(cs - mem\_amt)  
						\State mm.evict(cs - mem\_amt  - mm.free\_mem) 
						\State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem))
						\State sm.write(cs - mem\_amt)
					\EndIf

				\end{algorithmic}
			\end{algorithm}

			Algorithm~\ref{alg:write} describes our simulation of chunk write in 
			the IOController. 
			In the chunk write, data can be written to cache with memory bandwidth 
			without waiting for flushing or being throttled before the amount of 
			dirty data reaches dirty\_ratio, or otherwise written with disk bandwidth.
			Thus, our algorithm initially checks the remaining amount of data that 
			can be written as dirty data (line 6).
			If this amount is greater than 0, MemoryManager is asked to evict 
			data from cache to accommodate the written data (line 8). 
			If the amount to evict calculated and passed in the evict function is less than 0, 
			the eviction does nothing.
			After evicting cache, the actual amount of data that can be written to 
			page cache is calculated (line 9), a cache write is then simulated 
			with this amount through memory model and data written is added to 
			page cache LRU lists (line 10).
			After dirty\_ratio is reached, dirty data has to be flushed before new data 
			can be written to cache again. 
			If there is remaining chunk data that has not been written, we flush and 
			evict data from cache as much as possible so that the this data can be 
			written to cache (line 13-14). 
			Here, MemoryManager does not guarantee that the amount of flushed 
			and evicted data is enough for the write of remaining data. 
			In the case it is less than the required amount, the remaining data 
			being written to cache can be flushed and evicted right away. 
			Thus, the amount of remaining data written and kept in cache 
			after the write is limited by the amount free memory, and this amount 
			is added to cache (line 15). 
			Finally, the disk write is simulated with the remaining data amount (line 16), 
			and the chunk write algorithm finishes.

@@

			Since bandwidth sharing was not simulated, this prototype does not support 
			concurrency: it is limited to single-threaded pipelines running on systems 
			with a single-core CPU. We used it in a first validation of our simulation 
			model, against a real sequential pipeline running on a real system.
			The source code of our Python prototype is available at 
			\url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
